---
title: "Time series analysis BASIC"
author: "Suhail AK"
date: "20 April 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Time series analysis on airpassenger data set


```{r}
library(MASS)
library(tseries)
library(stats)
frequency(air)
air <- AirPassengers 
class(air)
start(air)
end(air)
time(air)
View(AirPassengers)

plot(air)
abline(reg=lm(air ~time(air)))
```


```{r}

stlval <- stl(air,s.window = "periodic")
plot(stlval)
```


```{r}
#to plot trendline
plot(aggregate(air,FUN = mean))
class(air)
```

```{r}
#for seasonality check use boxplot
boxplot(air~cycle(air))

```


```{r}
#now to make varience equal we use log function

station <- plot(diff(log(air)))

stationair <- diff(log(air))

#now to check if the data is Stationary or not (use adf testing)
adf.test(stationair,alternative = "stationary")


```

Once the data is stationary then apply time series model
like eg: ARIMA model
- ARIMA Model
AR I MA
AR- auto regression (p-value coeffecient)
I- integration (d value)
MA- moving average (q value)

to find q value using autocorrelation function

ie find the line before the inverted line index
 in this case from the graph the value of q is 1

```{r}
#using autocorrelation function to predict q value
acf(stationair)

```

similarly to find P coeffecient using partial autocorrealtion function
ie. from the graph the value of P coeffecient is 0 


also to find coeffecient d we have to check how many times we used differentiation(diff()) to make mean constant.
hence in this case we differentiated once hence d=1

```{r}
pacf(stationair)
```

Now applying ARIMA model
ie c(p,d,q) in this case (0,1,1)
in predict function fit is a model and n.ahead =(no of years)*frequency/period
and in pred1 converting the values from log to decimal form using e^

```{r}
fit <- arima(log(air),c(0,1,1),list(order=c(0,1,1),period=12))
pred <- predict(fit,n.ahead = 10*12)

pred1 <- 2.718^pred$pred

ts.plot(air,pred1,log="y",lty=c(1,3))

```

Now to test the model for the year 1960

```{r}

air1=ts(air,frequency = 12,start = c(1949,1),end = c(1959,12))
air1
fit1 <- arima(log(air1),c(0,1,1),seasonal=list(order=c(0,1,1),period=12))
pred2 <- predict(fit1,n.ahead=10*12)
pred2deci <- 2.718^pred2$pred
data1 <- head(pred2deci,12)
data1
pred_data1960 <- round(data1,0)
actual_data1960 <- tail(air,12)
df <- data.frame(pred_data1960,actual_data1960)
df
```


















